---
title: "Discussion - Week5"
author: "David Simbandumwe"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(matlib) 
```


# Example 1.2 (Coin Tossing) 

As we have noted, our intuition suggests that the probability of obtaining a head on a single toss of a coin is 1/2. To have the computer toss a coin, we can ask it to pick a random real number in the interval [0, 1] and test to see if this number is less than 1/2. If so, we shall call the outcome heads; if not we call it tails. Another way to proceed would be to ask the computer to pick a random integer from the set {0,1}. The program CoinTosses carries out the experiment of tossing a coin n times. Running this program, with n = 20, resulted in:


```{r}
n <- 20
result_lst <- vector(mode='list', length=n)

for (i in 1:n ) {
	result_lst[i] <- rbinom(1, 1, 0.5)
}

```


# Exercises 1

Modify the program CoinTosses to toss a coin n times and print out after every 100 tosses the proportion of heads minus 1/2. Do these numbers appear to approach 0 as n increases? Modify the program again to print out, every 100 times, both of the following quantities: the proportion of heads minus 1/2, and the number of heads minus half the number of tosses. Do these numbers appear to approach 0 as n increases?


`
```{r}

n <- 10000
calc_count = 1000

result_lst <- vector(mode='list', length=n)

for (i in 1:n ) {
	result_lst[i] <- rbinom(1, 1, 0.5)
	
	if (i %% calc_count == 0) {
		
		s <- sum(unlist(result_lst)) / i - 0.5
		h <- sum(unlist(result_lst)) - (0.5 * i)
		print(paste0(i, ' = ' , s , ' : ' , h ))
	}
	
}



```



