---
title: "DATA605_w15_Multivariable Calculus"
author: "David Simbandumwe"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
editor_options: 
  chunk_output_type: inline

---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
rm(list=ls())

```


```{r, include=FALSE}

library(matlib) 
library(ggplot2)
library(dplyr)

```



# Question 1
Find the equation of the regression line for the given points. Round any final values to the nearest hundredth, if necessary. (5.6,8.8), (6.3,12.4), (7,14.8), (7.7,18.2), (8.4,20.8)


```{r}

df <- data.frame(
	x = c(5.6,6.3,7,7.7,8.4),
	y = c(8.8,12.4,14.8,18.2,20.8)
)

df %>% ggplot() + geom_point(aes(x=x, y=y)) +
	xlim(0, 25) + ylim(0, 25)

lm <- lm(df, formula = y ~ x)
summary(lm)


```


$$ 
 y = a + \beta x \\
 y = -14.8 + 4.26x
$$



# Question 2
2. Find all local maxima, local minima, and saddle points for the function given below. Write your answer(s) in the form ( x, y, z ). Separate multiple points with a comma.


```{r}
d_dx = D(expr(24*x - 6*x*y^2 - 8*y^3),'x')
d_dy = D(expr(24*x - 6*x*y^2 - 8*y^3),'y')
d_dx
d_dy
```


\[f(x,y) = 24x - 6xy^2  - 8y^3 \\ \]
- first derivative x
\[
dx(x,y) = 6(y^2 - 4) \\
\text{set dx =  0} \\
0 = 6(y^2 - 4) \\
y = +- 2 \\
\text{-------------} \\
f(x,y) = 24x - 6xy^2  - 8y^3 \\
f(2,4) = 24x - 6xy^2  - 8y^3 \\
f(2,2) = -64  \\
(x,y,z) = (2,4,-64 ) \\
\]
- second derivative x
\[dx^2(x,y) = 0 \\
(x,y,z) = (4,2,-64 ) \text{ is a sadle point}
\]





- first derivative y
\[
dy(x,y) = -12xy + 24y^2 \\
\text{set dy =  0} \\
0 = -12xy + 24y^2 \\
12xy = 24y^2 \\
x = 2y \\
\text{set y = +-2} \\
x = +-4 \\
\text{-------------} \\
f(-2,-4) = 24x - 6xy^2  - 8y^3 \\
f(-2,-4) = 64  \\
(x,y,z) = (-4,-2,64) \text{ is a sadle point}\\
\]

- second derivative y
\[
dy^2(x,y) = -12x + 48y
\text{set dy =  0} \\
0 = -12x + 48y \\
12x = 48y
x = 2y
\text{set x = 0} \\
(x,y,z) = (-4,-2,64 ) \text{ is a sadle point}
\]



```{r}

x <- 4
y <- 2

f = expression(24*x - 6*x*y^2  - 8*y^3)
eval(f)


x <- -4
y <- -2

f = expression(24*x - 6*x*y^2  - 8*y^3)
eval(f)

```




# Question 3
3. A grocery store sells two brands of a product, the "house" brand and a "name" brand. The manager estimates that if she sells the "house" brand for x dollars and the "name" brand for y dollars, she will be able to sell 81 - 21x + 17y units of the "house" brand and 40 + 11x - 23y units of the "name" brand.



Step 1. Find the revenue function R ( x, y ).
\[R(x,y) = -21x^2+81x +28xy - 23y^2 + 40y\]

Step 2. What is the revenue if she sells the "house" brand for $2.30 and the "name" brand for $4.10?

```{r}
x <- 2.30
y <- 4.10
r <- expression(-21*x^2+81*x +28*x*y - 23*y^2 + 40*y)



print(paste0('total revenue = ', eval(r)))
```




# Question 4
4. A company has a plant in Los Angeles and a plant in Denver. The firm is committed to produce a total of 96 units of a
product each week. The total weekly cost is given by \[C(x, y) = \frac{1}{6}x^2 + \frac{1}{6} y^2 + 7x + 25y + 700\], where x is the number of units produced in Los Angeles and y is the number of units produced in Denver. How many units should be produced in each plant to minimize the total weekly cost?


```{r}

y <- 0:96
f <- function(y){return(1/6*(96-y)^2 + 1/6*y^2 + 7*(96-y) + 25*y +700)}
curve(f, from = 20, to = 22, type="l", main = 'Graphs', col = 'blue')

```



\[
C(x, y) = \frac{1}{6}x^2 + \frac{1}{6} y^2 + 7x + 25y + 700 \\
\text{ x = 96 - y} \\
C(x, y) = \frac{1}{6}(96-y)^2 + \frac{1}{6} y^2 + 7x + 25y + 700 \\
Cy(x,y) = \frac{2}{3} (y - 21)
\text{ xCy(x,y) = 0} \\
0 = \frac{2}{3} (y - 21)
y = 21
x = 96 - 21 = 75
C(75,21) = 2761
\]




```{r}

x <- 75
y <- 21
f <- function(x,y){return(1/6*x^2 + 1/6*y^2 + 7*x + 25*y +700)}
print(paste0('4) total cost C(75,21) =  ', f(75,21) ))

```




# Question 5
5. Evaluate the double integral on the given region.
\[\int\int (e^{8x+3y})  dA ;$  $R: 2 \le x \le 4$ and $2 \le y \le4\]


Write your answer in exact form without decimals.
\[
\int_{2}^{4}\int_{2}^{4} (e^{8x+3y}) dxdy \\
\text{from wolframalpha} \\
\frac{1}{24}(e^{22}-e^{28}-e^{38}+e^{44})
\]











