---
title: "DATA605_w12_Regression Analysis"
author: "David Simbandumwe"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
rm(list=ls())
options(scipen=999)

```


```{r, include=FALSE}

library(matlib) 
library(pracma)
library(readr)

library(explore)
library(graphics)
library(ggplot2)

library(ResourceSelection)
library(glmnet)

library(MASS)
library(tidyverse)
library(caret)
library(leaps)
library(regclass)

library(dplyr)


```



# Overview
The attached who.csv dataset contains real-world data from 2008. The variables included follow.
Country: name of the country
LifeExp: average life expectancy for the country in years
InfantSurvival: proportion of those surviving to one year or more
Under5Survival: proportion of those surviving to five years or more
TBFree: proportion of the population without TB.
PropMD: proportion of the population who are MDs
PropRN: proportion of the population who are RNs
PersExp: mean personal expenditures on healthcare in US dollars at average exchange rate
GovtExp: mean government expenditures per capita on healthcare, US dollars at average exchange rate 
TotExp: sum of personal and government expenditures.



```{r}

df <- read_csv('Homework/data/xid-528551256_1.csv')

```



# Question 1
Provide a scatterplot of LifeExp~TotExp, and run simple linear regression. Do not transform the variables. Provide and interpret the F statistics, R^2, standard error,and p-values only. Discuss whether the assumptions of simple linear regression met.

```{r}

ggplot(df, aes(x=TotExp, y=LifeExp)) + geom_point()
df_lm.1 <- lm(data = df, LifeExp ~ TotExp)
s_lm <- summary(df_lm.1)
s_lm

```




<b><i>

- F-statistic =`r round(s_lm$fstatistic['value'],4)`, corresponding p-value = `r round(s_lm$coefficients[2,4],4)`. Since this p-value is less than .05, the model as a whole is statistically significant.

- Multiple R-squared = `r round(s_lm$adj.r.squared,4)` This tells us that `r round(s_lm$adj.r.squared*100,2)`% of the variation in the response variable, LifeExp, can be explained by the predictor variable, TotExp.

Even thought the model is statistically significant as a whole it has limited predictive power. 

</b></i>



# Question 2
Raise life expectancy to the 4.6 power (i.e., LifeExp^4.6). Raise total expenditures to the 0.06 power (nearly a log transform, TotExp^.06). Plot LifeExp^4.6 as a function of TotExp^.06, and r re-run the simple regression model using the transformed variables. Provide and interpret the F statistics, R^2, standard error, and p-values. Which model is "better?"


```{r}

df$LifeExp2 <- df$LifeExp^4.6
df$TotExp2 <- df$TotExp^0.06

ggplot(df, aes(x=TotExp2, y=LifeExp2)) + geom_point()
df_lm.2 <- lm(data = df, LifeExp2 ~ TotExp2)
s_lm <- summary(df_lm.2)
s_lm


```


<b><i>

- F-statistic =`r round(s_lm$fstatistic['value'],4)`, corresponding p-value = `r round(s_lm$coefficients[2,4],4)`. Since this p-value is less than .05, the model as a whole is statistically significant.

- Multiple R-squared = `r round(s_lm$adj.r.squared,4)` This tells us that `r round(s_lm$adj.r.squared*100,2)`% of the variation in the response variable, LifeExp2, can be explained by the predictor variable, TotExp2.

The changes to the model have significantly increased the predictive power of the model while maintaining its statistical significance. 

</b></i>


# Question 3
Using the results from 3, forecast life expectancy when TotExp^.06 =1.5. Then forecast life expectancy when TotExp^.06=2.5.


```{r}

pred_df1 <- data.frame(TotExp2 = c(1.5,2.5))
pred_df1$LifeExp2 <- predict(df_lm.2,pred_df1)
pred_df1

```




# Question 4
Build the following multiple regression model and interpret the F Statistics, R^2, standard error, and p-values. How good is the model?

$$ LifeExp = b0+b1 x PropMd + b2 x TotExp +b3 x PropMD x TotExp $$
```{r}

df_lm.3 <- lm(data = df, LifeExp ~ PropMD + TotExp + PropMD * TotExp)
s_lm <- summary(df_lm.3)
s_lm



```

<b><i>

- F-statistic =`r round(s_lm$fstatistic['value'],4)`, corresponding p-value = `r round(s_lm$coefficients[2,4],4)`. Since this p-value is less than .05, the model as a whole is statistically significant.

- Multiple R-squared = `r round(s_lm$adj.r.squared,4)` This tells us that `r round(s_lm$adj.r.squared*100,2)`% of the variation in the response variable, LifeExp2, can be explained by the predictor variable, TotExp2 and PropPD. The adjusted R-squared highlights the limited explanitory power or the model. 

</b></i>



# Question 5
Forecast LifeExp when PropMD=.03 and TotExp = 14. Does this forecast seem realistic? Why or why not?


```{r}

pred_df2 <- data.frame(PropMD = .03, TotExp = 15)
pred_df2$LifeExp <- predict(df_lm.3,pred_df2)
pred_df2


```



<b><i>

- No it does not seem realistic. Although the PropMD is high 0.03 and would indicate a higher life expectancy, the TotExp is low at 15 indicating a lower life expectancy. 
- There are not many data points that exibit the characteristics of these two threshold. This limits the ability of the model to accuratly forcast values with these inputs.
- 

</b></i>



```{r}
df1 <- df %>% filter(PropMD < 0.03)

ggplot(df1, aes(x=PropMD, y=LifeExp)) + geom_point()
ggplot(df1, aes(x=TotExp, y=LifeExp)) + geom_point()
ggplot(df1, aes(x=TotExp, y=PropMD)) + geom_point()


```



```{r}

df1 <- df %>% filter(TotExp < 50)

ggplot(df1, aes(x=PropMD, y=LifeExp)) + geom_point()
ggplot(df1, aes(x=TotExp, y=LifeExp)) + geom_point()
ggplot(df1, aes(x=TotExp, y=PropMD)) + geom_point()

```




<br><br>
<b><i>Please show your work using an R-markdown document. Please name your assignment submission with your first initial and last name.</i></b>

