---
title: "DATA605_w10_Markov Chains and Random Walks"
author: "David Simbandumwe"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
rm(list=ls())

```


```{r, include=FALSE}

library(matlib) 
library(pracma)
library(stats)
library(scales)
library(expm)

```



# Question 1 
Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars.
A guard agrees to make a series of bets with him. If Smith bets A dollars, he wins A dollars with probability .4 and loses A dollars with probability .6.

Find the probability that he wins 8 dollars before losing all of his money if

Facts
- start 1 dollar needs 8 dollar
- win 1 dollar .4
- loss dollar .6



(a) he bets 1 dollar each time (timid strategy).

<b><i>The probablity that Smith gets out of jail with this strategy is 0.020</b></i>


```{r}

P <- matrix(
	c(0,.6,0,0,0,0,0,0,0,
	.4,0,.6,0,0,0,0,0,0,
	0,.4,0,.6,0,0,0,0,0,
	0,0,.4,0,.6,0,0,0,0,
	0,0,0,.4,0,.6,0,0,0,
	0,0,0,0,.4,0,.6,0,0,
	0,0,0,0,0,.4,0,0,0,
	.6,0,0,0,0,0,0,1,0,
	0,0,0,0,0,0,.4,0,1
	),9
)
P

z <- P%^%10000000
z <- round(z,3)
z
```




(b) he bets, each time, as much as possible but not more than necessary to bring his fortune up to 8 dollars (bold strategy).

<b><i>The probablity that Smith gets out of jail with this strategy is 0.064</b></i>


```{r}

P <- matrix(
		c(0,0,0,0,0,0,0,0,0,
		.4,0,0,0,.6,0,0,0,0,
		0,0,0,0,0,0,0,0,0,
		0,.4,0,0,0,.6,0,0,0,
		0,0,0,0,0,0,0,0,0,
		0,0,.4,0,0,0,.6,0,0,
		0,0,0,0,0,0,0,0,0,
		.6,.6,.6,.6,0,0,0,1,0,
		0,0,0,.4,.4,.4,.4,0,1
	),9
)
P

z <- P%^%1000000000
z <- round(z,3)
z
```


(c) Which strategy gives Smith the better chance of getting out of jail?

<b><i>The second strategy provides the best opportunity for success. Starting with 1 dollar the second strategy has a 0.064 probability of success vs 0.020 probability of success in the conservative betting approach</b></i>




g an R-markdown document. Please name your assignment submission with your first initial and last name.</i></b>

