---
title: "DATA605_w11_Regression Analysis"
author: "David Simbandumwe"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
rm(list=ls())

```


```{r, include=FALSE}

library(matlib) 
library(pracma)
library(stats)
library(scales)
library(expm)
library(dplyr)
library(explore)


```



# Question 1 
Using the “cars” dataset in R, build a linear model for stopping distance as a function of speed and replicate the analysis of your textbook chapter 3 (visualization, quality evaluation of the model, and residual analysis.)



```{r}
data(cars)
summary(cars)
cars %>% explore(speed, target = dist, split = TRUE)
cars %>% explain_tree(target = dist)

```

<b><i> 

- There appears to be a positive correlation between speed and distance
- The tree view highlights some clear breakpoints in speed that correspond to higher stopping distances

</i></b>





```{r}

car_lm <- lm(formula = dist ~ speed, data=cars)
s_lm <- summary(car_lm)
s_lm

```

<b><i> 

- The model explains `r s_lm$adj.r.squared` of the variability in stopping distance. 
- With a p value of `r s_lm$coefficients[2,4]` we can reject the null hypothesis in favor of the alternate hypothesis

</i></b>


```{r}
par(mfrow=c(2,2))
plot(car_lm)
```
<b><i> 

- The right side of the Q-Q plot diverges from the guide line, indicating that the residuals are not normally distributed

</i></b>


<br><br>
<b><i>Please show your work using an R-markdown document. Please name your assignment submission with your first initial and last name.</i></b>

