---
title: "DATA605_w13_Univariate and Multivariate Calculus"
author: "David Simbandumwe"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
rm(list=ls())
options(scipen=999)

```


```{r, include=FALSE}

library(matlib) 
library(ggplot2)

```



# Question 1
Use integration by substitution to solve the integral below.


$$

\int 4e^{-7x} dx \\

assume: u = -7x \\

\int \frac{4}{-7}e^{u} du \\
\frac{4}{-7}e^{u} + C \\

u = -7x \\

answer: 
- \frac{4}{7}e^{-7x} + C

$$





# Question 2
Biologists are treating a pond contaminated with bacteria. The level of contamination is changing at a rate of $\frac{dN}{dt} = - \frac{3150}{t^{4}} - 220$ bacteria per cubic centimeter per day, where t is the number of days since treatment began. Find a dt t4 function N( t ) to estimate the level of contamination if the level after 1 day was 6530 bacteria per cubic centimeter.

$$

\frac{dN}{dt} = - \frac{3150}{t^{4}} - 220 \\
t = 1, bacteria= 6530 cm^3 \\
f(1) = \int - \frac{3150}{t^{4}} - 220 \\
f(1) = \frac{3150}{3t^{3}} - 220t + C \\
f(1) = \frac{1050}{t^3} - 220t + C\\
t = 1, bacteria= 6530 cm^3 \\
f(1) = \frac{1050}{1^3} - 220(1) + C \\
f(1) = \frac{1050}{1^3} - 220(1) + C = 6530 cm^3\\
C = 5700 cm^3\\
f(1) = \frac{1050}{t^3} - 220t +  5700 \\

$$



# Question 3

Find the total area of the red rectangles in the figure below, where the equation of the line is f ( x ) = 2x - 9.


```{r}

f <- function(x){return(2*x - 9)}
curve(f, from = -5, to = 10, type="l", main = 'Graphs', col = 'blue')
area.f <- integrate(f, lower = 4, upper = 8)
print(paste0('3) area =  ', area.f$value ))

```




# Question 4
Find the area of the region bounded by the graphs of the given equations.
y = x2 - 2x - 2, y = x + 2
Enter your answer below.

```{r}

f <- function(x){return(x^2 - 2*x - 2)}
g <- function(x){return(x + 2)}
curve(f, from = -2, to = 5, type="l", main = 'Graphs', col = 'blue', ylim=c(-10,10))
curve(g, from = -2, to = 5, type="l", col = 'red', add=TRUE)

area.f <- integrate(f, lower = -1, upper = 4)
area.g <- integrate(g, lower = -1, upper = 4)
area = area.g$value - area.f$value


print(paste0('4) area =  ', round(area,4) ))


```




# Question 5
A beauty supply store expects to sell 110 flat irons during the next year. It costs \$3.75 to store one flat iron for one year. There is a fixed cost of \$8.25 for each order. Find the lot size and the number of orders per year that will minimize inventory costs.


```{r}

f <- function(n){return(3.75*(110/(2*n)) + 8.25*n)}
curve(f, from = 1, to = 40, type="l", main = 'Graphs', col = 'blue', xlim=c(1,40))

f(3)
f(4)
f(5)
f(6)
f(7)
f(8)
f(9)

```

$$
c = 3.75 * \frac{110}{2n} + 8.25n \\
\frac{dc}{dn} = \frac{206.25}{n} + 8.25n \\
\frac{dc}{dn} = -206.25n^-2 + 8.25 \\
0 = -206.25n^-2 + 8.25 \\
\frac{-8.25}{-206.25} = n^{-2}
5 = n

$$

5 orders in the year will reduce the overall costs.





<br><br>
<b><i>Please show your work using an R-markdown document. Please name your assignment submission with your first initial and last name.</i></b>

