---
title: "DATA605_w3_ Eigenvalues and Eigenvectors"
author: "David Simbandumwe"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
```


```{r}
library(matlib) 
library(pracma)
```


## Problem set 1


<b><i>(1) What is the rank of the matrix A?</i></b>

$A = \begin{bmatrix} 1&2&3&4 \\ -1&0&1&3 \\ 0&1&-2&1 \\5&4&-2&-3 \end{bmatrix}$

```{r}
A = matrix(c(1,-1,0,5, 2,0,1,4, 3,1,-2,-2, 4,3,1,-3),4)
A

print('- RREF')
rref(A)

print('- matrix rank')
Rank(A)
```

|		The rank of matrix A is = 4


<br><br><br>
<b><i>(2) Given an mxn matrix where m > n, what can be the maximum rank? The minimum rank, assuming that the matrix is non-zero?</i></b>

|		The maximum rank of the matrix is n and the minimum rank is 1 


<br><br><br>
<b><i>(3) What is the rank of matrix B?</i></b>

$B = \begin{bmatrix} 1&2&1 \\ 3&6&3 \\ 2&4&2 \end{bmatrix}$


```{r}
B = matrix(c(1,3,2, 2,6,4, 1,3,2),3)
B

#- RREF
rref(B)

#- matrix rank
Rank(B)


```

|		The rank of matrix A is = 1

<br><br><br>
<b><i>(4)Compute the eigenvalues and eigenvectors of the matrix A. You’ll need to show your
work. You’ll need to write out the characteristic polynomial and show your solution.</i></b>

$A = \begin{bmatrix} 1&2&3 \\ 0&4&5 \\ 0&0&6 \end{bmatrix}$

```{r}
A <- matrix(c(1,0,0, 2,4,0, 3,5,6), 3)
A

#eigen values / vectors'
e <- eigen(A)
print(e$values)

```


```{r}
#lamda 6
A1 <- 6*diag(3) - A
A1
rref(A1)


```





```{r}
#lamda 4
A2 <- 4*diag(3) - A
A2
rref(A2)


```




```{r}
#lamda 1
A3 <- 1*diag(3) - A
A3
rref(A3)
```





<br><br>
<b><i>Please show your work using an R-markdown document. Please name your assignment submission with your first initial and last name.</i></b>

