---
title: "DATA605_w3_ Eigenvalues and Eigenvectors"
author: "David Simbandumwe"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
```


```{r}
library(matlib) 
library(pracma)
```


## Problem set 1


<b><i>(1) What is the rank of the matrix A?</i></b>

$A = \begin{bmatrix} 1&2&3&4 \\ -1&0&1&3 \\ 0&1&-2&1 \\5&4&-2&-3 \end{bmatrix}$

```{r}
A = matrix(c(1,-1,0,5, 2,0,1,4, 3,1,-2,-2, 4,3,1,-3),4)
A

print('- RREF')
rref(A)

print('- matrix rank')
Rank(A)
```

|		The rank of matrix A is = 4


<br><br><br>
<b><i>(2) Given an mxn matrix where m > n, what can be the maximum rank? The minimum rank, assuming that the matrix is non-zero?</i></b>

|		The maximum rank of the matrix is n and the minimum rank is 1 


<br><br><br>
<b><i>(3) What is the rank of matrix B?</i></b>

$$B = \begin{bmatrix} 1&2&1 \\ 3&6&3 \\ 2&4&2 \end{bmatrix} $$
|		The 3 rows in the matrix are scaled equivalent there for the rank is 1 

```{r}
B = matrix(c(1,3,2, 2,6,4, 1,3,2),3)
B

#- RREF
rref(B)

#- matrix rank
Rank(B)


```

|		The rank of matrix A is = 1


<br><br>

## Problem Set 2

<br>		
<b><i>(4)Compute the eigenvalues and eigenvectors of the matrix A. You’ll need to show your
work. You’ll need to write out the characteristic polynomial and show your solution.</i></b>

$A = \begin{bmatrix} 1&2&3 \\ 0&4&5 \\ 0&0&6 \end{bmatrix}$


|		The eigenvalues of a triangular matrix are the diagonal 1,4,6. We can confirm that using R as follows

```{r}
A <- matrix(c(1,0,0, 2,4,0, 3,5,6), 3)
A

#eigen values / vectors'
e <- eigen(A)
print(e$values)

```

lambda 1

$$ 1\lambda \to (I - A) =  \begin{bmatrix} 1-1&-2&-3 \\ 0&1-4&-5 \\ 0&0&1-6 \end{bmatrix} =  \begin{bmatrix} 0&-2&-3 \\ 0&-3&-5 \\ 0&0&-5 \end{bmatrix} RREF \begin{bmatrix} 0&0&0 \\ 0&1		&0 \\ 0&0&1 \end{bmatrix} $$

- R3*(-1/5)
- R2 + (5)*R3
- R2*(-1/3)
- R1 + (-3)*R3 and R1 + (-2)*R2

$$ null space \to  1\lambda \to (I - A) = \begin{bmatrix} 0&0&0 \\ 0&1&0 \\ 0&0&1 \end{bmatrix} * \begin{bmatrix} x1 \\ x2 \\ x3 \end{bmatrix} = \begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix} $$

- x1=1, x2=0, x3=0

$$ N(I-A) \to span ( s * \begin{bmatrix} 1 \\ 0 \\ 0 \end{bmatrix}) $$


```{r}
#lambda 1
A1 <- 1*diag(3) - A
A1
rref(A1)

```


lambda 4

$$ 4\lambda \to (4I - A) =  \begin{bmatrix} 4-1&-2&-3 \\ 0&4-4&-5 \\ 0&0&4-6 \end{bmatrix} =  \begin{bmatrix} 3&-2&-3 \\ 0&0&-5 \\ 0&0&-2 \end{bmatrix} RREF \begin{bmatrix} 1&-2/3&0 \\ 0&0&0 \\ 0&0&1 \end{bmatrix} $$

- R3*(-1/2)
- R2 + (5)*R3 and R1 + (3)*R3
- R2*(1/3)

$$ null space \to  1\lambda \to (I - A) = \begin{bmatrix} 1&-2/3&0 \\ 0&0&0 \\ 0&0&1 \end{bmatrix} * \begin{bmatrix} x1 \\ x2 \\ x3 \end{bmatrix} = \begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix} $$

- x1=2/3, x2=1, x3=0

$$ N(I-A) \to span ( s * \begin{bmatrix} 2/3 \\ 1 \\ 0 \end{bmatrix}) $$




```{r}
#lambda 4
A2 <- 4*diag(3) - A
A2
rref(A2)

```


lambda 6

$$ 4\lambda \to (4I - A) =  \begin{bmatrix} 6-1&-2&-3 \\ 0&6-4&-5 \\ 0&0&6-6 \end{bmatrix} =  \begin{bmatrix} 5&-2&-3 \\ 0&2&-5 \\ 0&0&0 \end{bmatrix} RREF \begin{bmatrix} 1&0&-8/5 \\ 0&1&-5/2 \\ 0&0&0 \end{bmatrix} $$

- R2*(1/2)
- R1 + (-2)*R2
- R1*(1/5)

$$ null space \to  1\lambda \to (I - A) = \begin{bmatrix} 1&0&-8/5 \\ 0&1&-5/2 \\ 0&0&0 \end{bmatrix} * \begin{bmatrix} x1 \\ x2 \\ x3 \end{bmatrix} = \begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix} $$

- x1=8/5, x2=5/2, x3=1

$$ N(I-A) \to span ( s * \begin{bmatrix} 8/5 \\ 5/2 \\ 1 \end{bmatrix}) $$




```{r}
#lambda 6
A3 <- 6*diag(3) - A
A3
rref(A3)


```




```{r}

```





<br><br>
<b><i>Please show your work using an R-markdown document. Please name your assignment submission with your first initial and last name.</i></b>

