---
title: "DATA605_w8_Distributions and Densities / Expected Value and Variance"
author: "David Simbandumwe"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
rm(list=ls())

```


```{r, include=FALSE}

library(matlib) 
library(pracma)
library(stats)
library(scales)

```



11 A company buys 100 lightbulbs, each of which has an exponential lifetime of 1000 hours. What is the expected time for the first of these bulbs to burn out? (See Exercise 10.)

```{r}
mu <- 1000

print(paste0('11) expected lifetime 100 lightbulbs in independent so expected lifetime for each is = ', round(mu,4)))

```


14 Assume that X1 and X2 are independent random variables, each having an exponential density with parameter λ. Show that Z = X1 − X2 has density


$$

fZ(z) = \frac{1}{2} \lambda e ^ {-\lambda |z|}  \\
fZ(z) = \int_{-\infty}^{\infty}  fX1(z-x2) fX2(x2) dx2 \\
fZ(z) = \int_{-\infty}^{\infty}  fX1(z-x2) fX2(x2) dx2 \\
fZ(z) = \int_{0}^{z} \lambda e ^ {-\lambda (x2 -x1)} \lambda e ^ {-\lambda x2} \\
fZ(z) = \int_{0}^{z} \lambda^2 e ^ {-\lambda x2 +\lambda x1 -\lambda x2} \\
fZ(z) = \int_{0}^{z} \lambda^2 e ^ {\lambda x1 -2\lambda x2 } \\
fZ(z) = \int_{0}^{z} \lambda^2 e ^ {-\lambda (x1 - x2) } \\
fZ(z) = \frac{1}{2} \lambda e ^ {-\lambda |z| } \\

$$






1. Let X be a continuous random variable with mean μ = 10 and variance σ2 = 100/3. Using Chebyshev’s Inequality, find an upper bound for the following probabilities.
 
(a) P(|X−10|≥2)
(b) P(|X−10|≥5)
(c) P(|X−10|≥9)
(d) P(|X − 10| ≥ 20)

$$ P(|x - \mu| \le a) =  (\frac{Var(x)}{a^2}) $$ 

```{r}

var <- 100/3


# (a) P(|X−10|≥2)
a <- 2
p2 <- var / a^2
print(paste0('1.a) probability = ', round(p2,4)))


#(b) P(|X−10|≥5)
a <- 5
p5 <- var / a^2
print(paste0('1.b) probability = ', round(p5,4)))


#(c) P(|X−10|≥9)
a <- 9
p9 <- var / a^2
print(paste0('1.c) probability = ', round(p9,4)))


#(d) P(|X − 10| ≥ 20)
a <- 20
p20 <- var / a^2
print(paste0('1.d) probability = ', round(p20,4)))

```




<br><br>
<b><i>Please show your work using an R-markdown document. Please name your assignment submission with your first initial and last name.</i></b>

