---
title: "DATA605_w7_Distributions and Densities / Expected Value and Variance"
author: "David Simbandumwe"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
rm(list=ls())

```


```{r, include=FALSE}

library(matlib) 
library(pracma)
library(stats)
library(scales)

```


1. Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi’s. Find the distribution of Y .

- 1 to k
- y min value of Xi

<br><br>
<b>
probability = $prob = \frac{1}{k}$

distribution function = $df = \frac{\left\lvert {k} \right\rvert - a + 1}{n}$
</b>

<br><br>
2. Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has a manufacturer’s expected lifetime of 10 years. This means that we expect one failure every ten years. (Include the probability statements and R Code for each part.).

- life = 10 years
- mean failure = 10 years
- 10 years between events


a. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a geometric. (Hint: the probability is equivalent to not failing during the first 8 years..)

- prob machine will fail in 8 years

```{r}

p <- 1/10
n <- 8
e <- 1/p
std <- sqrt((1-p)/p^2)

x_lst <- c(0,1,2,3,4,5,6,7,8)
prob <- (1-p)^(x_lst) * (p)
prob <- 1 - sum(prob)

print(paste0('2.d) probability = ', round(prob,4)))
print(pgeom(n,p, lower.tail = FALSE))
print(paste0('2.d) expected = ', round(e,4)))
print(paste0('2.d) std deviation = ', round(std,4)))

```




b. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as an exponential.


```{r}

mu <- 10
l <- 1/10
k <- 8

prob <- exp(-k/mu)
std <- mu

print(paste0('2.d) probability = ', round(prob,4)))
print(pexp(k, l, lower.tail = FALSE))
print(paste0('2.d) mean / expected = ', round(mu,4)))
print(paste0('2.d) std deviation = ', round(std,4)))

```


c. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8 years)


```{r}
p <- 1/10
n <- 8
x <- 0
e <- n*p
std <- sqrt(n*p*(1-p))

prob <- p^x * (1-p)^(n-x)
print(paste0('2.d) probability = ', round(prob,4)))
print(dbinom(0,8,0.1))

print(paste0('2.d) expected = ', round(e,4)))
print(paste0('2.d) std deviation = ', round(std,4)))

```




d. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a Poisson.


```{r}
r <- 1/10
t <- 8
l <- r*t
e <- l
x <- 0
std <- sqrt(l)

prob <- ((l^x) * exp(-l)) / factorial(x)

print(paste0('2.d) probability = ', round(prob,4)))
print(ppois(x, l))
print(paste0('2.d) expected = ', round(e,4)))
print(paste0('2.d) std deviation = ', round(std,4)))
```



<br><br>
<b><i>Please show your work using an R-markdown document. Please name your assignment submission with your first initial and last name.</i></b>

