---
title: "Scratch 4"
output: html_document
date: "2022-09-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(matlib) 
library(pracma)

```


```{r, include=FALSE}

library(matlib) 
library(pracma)
library(stats)
library(scales)

```


1. Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi’s. Find the distribution of Y .

- 1 to k
- y min value of Xi



```{r}

d <- '1/k'
print(paste0('1) distribution = ', d))

```


2. Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has a manufacturer’s expected lifetime of 10 years. This means that we expect one failure every ten years. (Include the probability statements and R Code for each part.).

- life = 10 years
- mean failure = 10 years
- 10 years between events


a. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a geometric. (Hint: the probability is equivalent to not failing during the first 8 years..)

- prob machine will fail in 8 years

```{r}

p <- 1/10
n <- 8
prob <- (1-p)^(n-1) * (p)
e <- 1/p
std <- sqrt((1-p)/p^2)
print(paste0('2.d) probability = ', round(prob,4)))
print(paste0('2.d) mean / expected = ', round(e,4)))
print(paste0('2.d) std deviation = ', round(std,4)))

```


```{r}
k <- 8
p <- 1/10

dgeom(k,p)
```



b. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as an exponential.


```{r}

mu <- 10
l <- 1/10
k <- 8

prob <- exp(-k/mu)
std <- mu

print(paste0('2.d) probability = ', round(prob,4)))
print(pexp(k, l, lower.tail = FALSE))
print(paste0('2.d) mean / expected = ', round(mu,4)))
print(paste0('2.d) std deviation = ', round(std,4)))

```


```{r}
l <- 1/20
mu <- 1/l
x <- 10

prob <- 1 - exp(-x/mu)
std <- mu
print(paste0('2.d) probability = ', round(prob,4)))
print(pexp(10, 1/20, lower.tail = TRUE))
print(paste0('2.d) mean / expected = ', round(mu,4)))
print(paste0('2.d) std deviation = ', round(std,4)))
```

```{r}

l <- 1/20
mu <- 1/l
x <- 30

prob <- exp(-x/mu)
std <- mu
print(paste0('2.d) probability = ', round(prob,4)))
print(pexp(30, 1/20, lower.tail = FALSE))
print(paste0('2.d) mean / expected = ', round(mu,4)))
print(paste0('2.d) std deviation = ', round(std,4)))

```


c. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8 years)



```{r}

p <- 1/10
n <- 8
x <- 0
e <- n*p
std <- sqrt(p*(1-p)/n)

prob <- p^x * (1-p)^(n-x)
print(paste0('2.d) probability = ', round(prob,4)))
print(dbinom(0,8,0.1))

print(paste0('2.d) expected = ', round(e,4)))
print(paste0('2.d) std deviation = ', round(std,4)))

```


d. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a Poisson.


```{r}
#mu <- 10
l <- 10
e <- l
x <- 8

prob <-((l^x) * exp(-l)) / factorial(x)

std <- sqrt(l)
print(paste0('2.d) probability = ', round(prob,4)))
print(ppois(x, l, lower.tail = FALSE))
print(paste0('2.d) expected = ', round(e,4)))
print(paste0('2.d) std deviation = ', round(std,4)))

```


```{r}

l <- 10
x <- 8
x_lst <- c(1:x)

p <- prob <-((l^x_lst) * exp(-l)) / factorial(x_lst)
prob <- 1 - sum(p)

print(paste0('2.d) probability = ', round(prob,4)))
print(ppois(x, l, lower.tail = FALSE))

```


```{r}

l <- 12
x <- 10
prob1 <- ((l^x) * exp(-l)) / factorial(x)
print(paste0('a) probability = ', round(prob1,4)))

x_lst <- c(1:x-1)
p <- ((l^x_lst) * exp(-l)) / factorial(x_lst)
prob2 <- 1 - sum(p)
print(paste0('b) probability = ', round(prob2,4)))

l <- 12/24
x <- 1
x_lst <- c(0,1)
p <- ((l^x_lst) * exp(-l)) / factorial(x_lst)
prob3 <- 1 - sum(p)
print(paste0('c) probability = ', round(prob3,7)))

```



- average rate of 3 per hour
- 1/6

```{r}

l <- 3
x <- 5

prob <-((l^x) * exp(-l)) / factorial(x)
mu <- 1/l
std <- sqrt(l)
print(paste0('2.d) probability = ', round(prob,4)))
print(dpois(x, l))
print(paste0('2.d) mean / expected = ', round(mu,4)))
print(paste0('2.d) std deviation = ', round(std,4)))


```


```{r}


l <- 3
x <- 5

prob <-((l^x) * exp(-l)) / factorial(x)
mu <- 1/l
std <- sqrt(l)
print(paste0('2.d) probability = ', round(prob,4)))
print(ppois(x, l, lower.tail = TRUE))
print(paste0('2.d) mean / expected = ', round(mu,4)))
print(paste0('2.d) std deviation = ', round(std,4)))


```







```{r}

x <- seq(0, 6, length=7)
y <- dunif(x, min = 0, max = 4)

```


