---
title: "DATA605_Final"
author: "David Simbandumwe"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
editor_options: 
  chunk_output_type: inline

---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
rm(list=ls())

```


```{r, include=FALSE}

library(matlib) 
library(pracma)
library(ggplot2)

library(Matrix)
library(igraph)
library(openintro)
library(expm) 
library(matrixcalc)

library(igraph)


library(tidyverse)
library(keras)
library(yardstick)
library(tidymodels)
library(workflows)

library(nnet)
library(caret)


library(corrplot)
library(GGally)
library(stats)
library(MASS)


library(textrecipes)
library(recipes)
library(modeldata)


```






# Problem 3: Advanced Regression Techniques

Modeling. Build some type of multiple regression model and submit your model to the competition board. Provide your complete model summary and results with analysis. Report your Kaggle.com user name and score. 10 points






## Load data

```{r}

# read in training data
house_train_df <- read_csv("./Final/house-prices/train.csv")
house_train_df %>% select_if(is.numeric) %>% summary()

# reokace b/a with 0 where applicable
house_train_df <- house_train_df %>% mutate(
                    LotFrontage = ifelse(is.na(LotFrontage), 0, LotFrontage),
                    MasVnrArea = ifelse(is.na(MasVnrArea), 0, MasVnrArea),
                    GarageYrBlt = ifelse(is.na(MasVnrArea), mean(GarageYrBlt, na.rm = TRUE), MasVnrArea),
                    FireplaceQu = ifelse(is.na(FireplaceQu), "None", FireplaceQu),
                    BsmtFinSF1 = ifelse(is.na(BsmtFinSF1), 0, BsmtFinSF1),
                    TotalBsmtSF = ifelse(is.na(TotalBsmtSF), 0, TotalBsmtSF),
                    PoolQC = ifelse(is.na(PoolQC), "None", PoolQC),
                    Street = ifelse(is.na(Street), "None", Street),
                    Alley = ifelse(is.na(Alley), "None", Alley),
                    GarageType = ifelse(is.na(GarageType), "None", GarageType),
                    GarageFinish = ifelse(is.na(GarageFinish), "None", GarageFinish),
                    GarageQual = ifelse(is.na(GarageQual), "None", GarageQual),
                    GarageCond = ifelse(is.na(GarageCond), "None", GarageCond),
                    BsmtQual = ifelse(is.na(BsmtQual), "None", BsmtQual),
                    KitchenQual = ifelse(is.na(KitchenQual), "None", KitchenQual),
                    GrLivArea = ifelse(is.na(GrLivArea), 0, GrLivArea)
                  )

#Print number of null
print('---- nulls ----')
all_nas <- colSums(sapply(house_train_df[names(house_train_df)], is.na))
all_nas[which(all_nas > 0)]


```



## Baseline - Full Numeric Model
This model will include all numeric fields as a baseline. 

```{r}

# moduel with numeric variables only
houseNum_df <- house_train_df %>%  dplyr::select_if(is.numeric)
ml.num <- lm(SalePrice ~., houseNum_df)


# stepwize optimizaiton
ml.num.step <- stats::step(ml.num, criteria='AIC', trace=FALSE,direction='both')
summary(ml.num.step)


# residuals
par(mfrow=c(2,2))
plot(ml.num.step)


```

lm(formula = SalePrice ~ MSSubClass + LotArea + OverallQual + 
    OverallCond + YearBuilt + YearRemodAdd + MasVnrArea + BsmtFinSF1 + 
    BsmtUnfSF + `1stFlrSF` + `2ndFlrSF` + LowQualFinSF + BsmtFullBath + 
    FullBath + BedroomAbvGr + KitchenAbvGr + TotRmsAbvGrd + Fireplaces + 
    GarageCars + WoodDeckSF + ScreenPorch, data = houseNum_df)
    

## Analysis - Candidate Variable

```{r}

candidates <- c('SalePrice','MSSubClass','LotArea','LotFrontage','OverallQual','YearBuilt','YearRemodAdd',
				'BsmtFinSF1','BsmtUnfSF','TotalBsmtSF','1stFlrSF','2ndFlrSF','GrLivArea','BsmtFullBath',
				'FullBath','BedroomAbvGr','KitchenAbvGr','TotRmsAbvGrd','Fireplaces','GarageCars','GarageArea',
				'WoodDeckSF','OpenPorchSF','FireplaceQu','SaleCondition','OverallCond','CentralAir','PoolQC',
				'ExterCond','MasVnrArea','TotalBsmtSF','ScreenPorch')

full_df <- house_train_df %>%
   dplyr::select(all_of(candidates)) %>%
   mutate(ExterCond=as.numeric(as.factor(ExterCond)),
			FullBath=as.numeric(as.factor(FullBath)),
			FireplaceQu=as.numeric(as.factor(FireplaceQu)), 
			SaleCondition=as.numeric(as.factor(SaleCondition)), 
			GarageCars=as.numeric(as.factor(GarageCars)),
			OverallCond=as.numeric(as.factor(OverallCond)), 
			CentralAir=as.numeric(as.factor(CentralAir)),
			PoolQC=as.numeric(as.factor(PoolQC))
   )

ames_corr <- full_df
ames_correlations <- cor(ames_corr, method = c("pearson"))
corrplot(ames_correlations, method="color")

```


```{r}

t <- ames_correlations
t[t < 0.3 ] <- 0
corrplot(t, method="color")


t <- as.data.frame(ames_correlations)
t <- t[t$SalePrice > 0.3]
t <- t %>% dplyr::filter(SalePrice > 0.3)


candidates <- names(t)
candidates

ames_corr %>% 
	dplyr::select(all_of(candidates)) %>% 
	ggpairs()

corrplot(as.matrix(t), method="color")


```







```{r}


ggplot(house_train_df, aes(y=SalePrice, x=OverallQual)) + 
	scale_y_continuous(labels = comma) + geom_point()

ggplot(house_train_df, aes(y=SalePrice, x=YearBuilt)) + 
	scale_y_continuous(labels = comma) + geom_point()

ggplot(house_train_df, aes(y=SalePrice, x=YearRemodAdd)) + 
	scale_y_continuous(labels = comma) + geom_point()

ggplot(house_train_df, aes(y=SalePrice, x=BsmtFinSF1)) + 
	scale_y_continuous(labels = comma) + geom_point()

ggplot(house_train_df, aes(y=SalePrice, x=TotalBsmtSF)) + 
	scale_y_continuous(labels = comma) + geom_point()

ggplot(house_train_df, aes(y=SalePrice, x=house_train_df$'1stFlrSF')) + 
	scale_y_continuous(labels = comma) + geom_point()

ggplot(house_train_df, aes(y=SalePrice, x=house_train_df$'2ndFlrSF')) + 
	scale_y_continuous(labels = comma) + geom_point()

ggplot(house_train_df, aes(y=SalePrice, x=GrLivArea)) + 
	scale_y_continuous(labels = comma) + geom_point()

ggplot(house_train_df, aes(y=SalePrice, x=FullBath)) + 
	scale_y_continuous(labels = comma) + geom_point()

ggplot(house_train_df, aes(y=SalePrice, x=TotRmsAbvGrd)) + 
	scale_y_continuous(labels = comma) + geom_point()

ggplot(house_train_df, aes(y=SalePrice, x=Fireplaces)) + 
	scale_y_continuous(labels = comma) + geom_point()

ggplot(house_train_df, aes(y=SalePrice, x=GarageCars)) + 
	scale_y_continuous(labels = comma) + geom_point()

ggplot(house_train_df, aes(y=SalePrice, x=GarageArea)) + 
	scale_y_continuous(labels = comma) + geom_point()

ggplot(house_train_df, aes(y=SalePrice, x=WoodDeckSF)) + 
	scale_y_continuous(labels = comma) + geom_point()

ggplot(house_train_df, aes(y=SalePrice, x=OpenPorchSF)) + 
	scale_y_continuous(labels = comma) + geom_point()

ggplot(house_train_df, aes(y=SalePrice, x=MasVnrArea)) + 
	scale_y_continuous(labels = comma) + geom_point()



```




```{r}


ggplot(house_train_df, aes(x=log(SalePrice))) + geom_histogram()
ggplot(house_train_df, aes(x=YearBuilt-min(house_train_df$YearBuilt))) + geom_histogram()
ggplot(house_train_df, aes(x=YearRemodAdd-min(house_train_df$YearRemodAdd))) + geom_histogram()
ggplot(house_train_df, aes(x=log(TotalBsmtSF))) + geom_histogram()
ggplot(house_train_df, aes(x=house_train_df$'1stFlrSF')) + geom_histogram()
ggplot(house_train_df, aes(x=GrLivArea)) + geom_histogram()
ggplot(house_train_df, aes(x=GarageArea)) + geom_histogram()


```





```{r}


ggplot(house_train_df, aes(y=SalePrice, x=YearBuilt)) + 
	scale_y_continuous(labels = comma) + geom_point() +
	facet_wrap(vars(Neighborhood))

ggplot(house_train_df, aes(y=SalePrice, x=YearBuilt)) + 
	scale_x_log10() + geom_point()
	
ggplot(house_train_df, aes(y=SalePrice, x=TotalBsmtSF)) + 
	scale_x_log10() + geom_point()

ggplot(house_train_df, aes(y=SalePrice, x=TotalBsmtSF)) + 
	scale_x_sqrt() + geom_point()


```



[1] "SalePrice"    "OverallQual"  "YearBuilt"    "YearRemodAdd" "BsmtFinSF1"   "TotalBsmtSF"  "1stFlrSF"     "2ndFlrSF"    
 [9] "GrLivArea"    "FullBath"     "TotRmsAbvGrd" "Fireplaces"   "GarageCars"   "GarageArea"   "WoodDeckSF"   "OpenPorchSF" 
[17] "MasVnrArea"  



## Full Model

```{r}


candidates <- c('SalePrice','OverallQual','YearBuilt','YearRemodAdd','BsmtFinSF1','BsmtUnfSF',
				'TotalBsmtSF',
				'1stFlrSF','2ndFlrSF','GrLivArea','BsmtFullBath','FullBath','HalfBath','KitchenAbvGr',
				'TotRmsAbvGrd','Fireplaces','GarageCars','GarageArea','WoodDeckSF','OpenPorchSF',
				'FireplaceQu',
				'SaleCondition','OverallCond','CentralAir','PoolQC','ExterCond','YearRemodAdd',
				'MasVnrArea',
				'MSSubClass','MSZoning','Street','Alley','LotShape','LandContour','LotConfig','HouseStyle',
				'RoofStyle','RoofMatl','BsmtQual','BsmtCond','BsmtFinType1','KitchenQual','GarageType',
				'GarageQual')


full_df <- house_train_df %>%
   dplyr::select(all_of(candidates)) %>%
   mutate(ExterCond=(as.factor(ExterCond)),
			MSSubClass=(as.factor(MSSubClass)),
			MSZoning=(as.factor(MSZoning)),
			Street=(as.factor(Street)),
			Alley=(as.factor(Alley)),
			LotShape=(as.factor(LotShape)),
			LandContour=(as.factor(LandContour)),
			LotConfig=(as.factor(LotConfig)),
			HouseStyle=(as.factor(HouseStyle)),
			RoofStyle=(as.factor(RoofStyle)),
			RoofMatl=(as.factor(RoofMatl)),
			BsmtQual=(as.factor(BsmtQual)),
			BsmtCond=(as.factor(BsmtCond)),
			BsmtFinType1=(as.factor(BsmtFinType1)),
			KitchenQual=(as.factor(KitchenQual)),
			FireplaceQu=(as.factor(FireplaceQu)), 
			GarageType=(as.factor(GarageType)), 
			GarageQual=(as.factor(GarageQual)), 
			SaleCondition=(as.factor(SaleCondition)), 
			OverallCond=c(as.factor(OverallCond)), 
			CentralAir=(as.factor(CentralAir)),
			PoolQC=(as.factor(PoolQC))
   )


#full_df <- ames_corr 
ml.full <- lm(formula = SalePrice ~ ., data = full_df)


# stepwize optimizaiton
ml.full.step <- stats::step(ml.full, criteria='AIC', trace=FALSE,direction='both')
summary(ml.full.step)


# residuals
par(mfrow=c(2,2))
plot(ml.full.step)

```




## Model 1

Removed
- GarageCars
- WoodDeckSF
- GarageArea
- OverallCond
- MSSubClass
- MSZoning
- BsmtFinType1
- LotConfig


```{r}


candidates <- c('SalePrice','OverallQual','YearBuilt','BsmtUnfSF','TotalBsmtSF',
				'1stFlrSF','2ndFlrSF','KitchenAbvGr', 'TotRmsAbvGrd','Fireplaces',
				'SaleCondition',
				'PoolQC', 'MasVnrArea','LotShape', 'LandContour',
				'RoofMatl', 'BsmtQual','KitchenQual')

df1 <- full_df %>% dplyr::select(all_of(candidates))


# model v1
ml.v1 <- lm(formula = SalePrice ~ ., data = df1)


# stepwize optimization
ml.v1.step <- stats::step(ml.v1, criteria='AIC', trace=FALSE,direction='both')
summary(ml.v1.step)


# residuals
par(mfrow=c(2,2))
plot(ml.v1.step)


```




## Predict
Selected model 2 for the predictions


```{r}

house_test_df <- read_csv("./Final/house-prices/test.csv")


# reokace b/a with 0 where applicable
house_test_df <- house_test_df %>% mutate(
                    LotFrontage = ifelse(is.na(LotFrontage), 0, LotFrontage),
                    MasVnrArea = ifelse(is.na(MasVnrArea), 0, MasVnrArea),
                    GarageYrBlt = ifelse(is.na(MasVnrArea), mean(GarageYrBlt, na.rm = TRUE), MasVnrArea),
                    FireplaceQu = ifelse(is.na(FireplaceQu), "None", FireplaceQu),
                    BsmtFinSF1 = ifelse(is.na(BsmtFinSF1), 0, BsmtFinSF1),
                    TotalBsmtSF = ifelse(is.na(TotalBsmtSF), 0, TotalBsmtSF),
                    PoolQC = ifelse(is.na(PoolQC), "None", PoolQC),
                    Street = ifelse(is.na(Street), "None", Street),
                    Alley = ifelse(is.na(Alley), "None", Alley),
                    GarageType = ifelse(is.na(GarageType), "None", GarageType),
                    GarageFinish = ifelse(is.na(GarageFinish), "None", GarageFinish),
                    GarageQual = ifelse(is.na(GarageQual), "None", GarageQual),
                    GarageCond = ifelse(is.na(GarageCond), "None", GarageCond),
                    BsmtQual = ifelse(is.na(BsmtQual), "None", BsmtQual),
                    KitchenQual = ifelse(is.na(KitchenQual), "Gd", KitchenQual),
                    GrLivArea = ifelse(is.na(GrLivArea), 0, GrLivArea),
                    BsmtUnfSF = ifelse(is.na(BsmtUnfSF), 0, BsmtUnfSF)
                  )



# identify variables with null values
all_nas <- colSums(sapply(house_test_df[names(house_test_df)], is.na))
all_nas[which(all_nas > 0)]




house_test_df <- house_test_df %>%
   mutate(ExterCond=(as.factor(ExterCond)),
			MSSubClass=(as.factor(MSSubClass)),
			MSZoning=(as.factor(MSZoning)),
			Street=(as.factor(Street)),
			Alley=(as.factor(Alley)),
			LotShape=(as.factor(LotShape)),
			LandContour=(as.factor(LandContour)),
			LotConfig=(as.factor(LotConfig)),
			HouseStyle=(as.factor(HouseStyle)),
			RoofStyle=(as.factor(RoofStyle)),
			RoofMatl=(as.factor(RoofMatl)),
			BsmtQual=(as.factor(BsmtQual)),
			BsmtCond=(as.factor(BsmtCond)),
			BsmtFinType1=(as.factor(BsmtFinType1)),
			KitchenQual=(as.factor(KitchenQual)),
			FireplaceQu=(as.factor(FireplaceQu)), 
			GarageType=(as.factor(GarageType)), 
			GarageQual=(as.factor(GarageQual)), 
			SaleCondition=(as.factor(SaleCondition)), 
			OverallCond=c(as.factor(OverallCond)), 
			CentralAir=(as.factor(CentralAir)),
			PoolQC=(as.factor(PoolQC))
   )


# predict new values
house_test_df$SalePrice <- predict.lm(ml.v1.step, house_test_df, type = "response")


# plot housing prices vs neighborhood
ggplot(house_test_df, aes(x=SalePrice, fill=Neighborhood)) + geom_histogram()

```




```{r}

# create results dataframe
result_df <- house_test_df %>% dplyr::select(Id,SalePrice)

# identify na predictions
all_nas <- colSums(sapply(result_df[names(result_df)], is.na))
all_nas[which(all_nas > 0)]

# write results
write.csv(result_df,"./Final/house-prices/result.csv", row.names = FALSE)


```







